Anomalous waves triggered by abrupt depth changes: experimental measurements and deterministic-statistical truncated KdV theory

### 2019-10-07 paper2_JNLS/paper.tex
NM

	Bolles et al.~systematically explored the effect of varying the driving amplitude on the anomalous statistics for a fixed depth ratio. Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we also report experiments in which the depth ratio was varied. To have a simple baseline for comparison against theory, we refer to the experiments from Fig.~\ref{ExpDiagStats} as the {\em reference} experiments. In these experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the cases $\Dphi = 1.25^{\circ}$ and $1.5^{\circ}$ shown in Fig.~\ref{ExpDiagStats}(a)), the upstream depth is $\dup = 12.5$ cm, and the downstream depth is $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$. From the data represented in Fig.\ref{ExpSpatialStats}, we extract a characteristic wave amplitude of $\etastd = 0.21$ cm for these representative experiments, which will be important for setting dimensionless parameters that enter the theory.

% NOTES
% Maybe modify slightly this last paragraph if it does not seem to fit next time I read it.
 %$\epsup = 0.017$.

@@

We now introduce the theoretical framework that will be used to understand and quantify the experimental observations. This framework is based on a Galerkin truncation of the variable-depth Korteweg–de Vries (KdV) equation. The KdV equation is a commonly excepted model for describing the propagation of unidirectional, shallow-water waves, accounting for weak nonlinearity and weak dispersion over long timescales and large spatial scales. We will perform Galerkin truncation of this system at wavenumber $\Lambda = $8--32 to obtain a finite dimensional dynamical system exhibiting weak turbulence. We outline the Hamiltonian structure of both the continuous KdV and the truncated system. This structure is exploited to obtain invariant measures of the underlying dynamics and to ultimately understand the experimental findings on anomalous wave statistics triggered by an ADC.

@@

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}[h]%[htbp]
\begin{center}
\caption{Table of parameters}
\label{paramtable}
\begin{tabular}{l l l}
\hline \multicolumn{3} { c }{Parameters that are constant in a single experiment} \\
\hline Description & Notation and definition & Value in experiments \\
\hline
Peak forcing frequency		& $f_p$						& 2 Hz \\
Characteristic wave amplitude	& $\etastd = \mean{\eta^2}^{1/2} $		& 0.03--0.3 cm \\
Upstream depth			& $\dup$						& 12.5 cm \\
Downstream depth			& $\ddn$						& 3 cm (and varied) \\
Upstream wavelength		& $\lamup = \sqrt{g \dup}/f_p$		& 55 cm \\
Upstream wavelength		& $\lamdn = \sqrt{g \ddn}/f_p$		& 27 cm \\
%
Amplitude-to-depth ratio		& $\epsup = \etastd / \dup$			& 0.0024--0.024 \\
Depth-to-wavelength ratio		& $\delup = \dup / \lamup$		& 0.22 \\
Depth ratio				& $\dratdn = \ddn/\dup$			& 0.24 (and varied)
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

>>>>

	Bolles et al.~systematically explored the effect of varying the driving amplitude on the anomalous statistics for a fixed depth ratio. Here, we further explore the effect of driving amplitude, especially in conjunction with theoretical developments, and we also report experiments in which the depth ratio was varied. To have a simple baseline for comparison against theory, we refer to the data from Fig.~\ref{ExpDiagStats} as the {\em reference} experiments. In these experiments, the driving amplitude is $\Dphi = 1.38^{\circ}$ (intermediate between the cases $\Dphi = 1.25^{\circ}$ and $1.5^{\circ}$ shown in Fig.~\ref{ExpDiagStats}(a)), the upstream depth is $\dup = 12.5$ cm, and the downstream depth is $\ddn = 3$ cm, giving a depth ratio of $\dratdn = \ddn/\dup= 0.24$. From the data represented in Fig.~\ref{ExpSpatialStats}, we extract a characteristic wave amplitude for the representative experiments of $\etastd = 0.21$ cm, which will be important for setting dimensionless parameters that enter the theory. Table \ref{paramtable} lists the range of experimental parameters along with these representative values, as well as value of dimensionless parameters that will be introduced later.

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}[h]%[htbp]
\begin{center}
\caption{Table of parameters}
\label{paramtable}
\begin{tabular}{l l l l}
%\hline \multicolumn{3} { c }{Parameters that are constant in a single experiment} \\
\hline Description & Notation & Experimental range & `Representative' values \\
\hline
Peak forcing frequency	& $f_p$					& 2 Hz 		& 2 Hz \\
Characteristic amplitude	& $\etastd$				& 0.03--0.3 cm	& 0.21 cm \\
Upstream depth		& $\dup$					& 12.5 cm 	& 12.5 cm \\
Downstream depth		& $\ddn$					& RANGE		& 3 cm \\
Upstream wavelength	& $\lamup = \sqrt{g \dup}/f_p$	& 55 cm 		& 55 cm \\
Downstream wavelength	& $\lamdn = \sqrt{g \ddn}/f_p$	& RANGE		& 27 cm \\
%
Amplitude-to-depth ratio	& $\epsup = \etastd / \dup$	&0.0024--0.024	& 0.017 \\
Depth-to-wavelength ratio	& $\delup = \dup / \lamup$	& RANGE		& 0.22 \\
Depth ratio			& $\dratdn = \ddn/\dup$		& RANGE		& 0.24 \\
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

% NOTES
% Maybe modify slightly the last paragraph if it does not seem to fit next time I read it.
 %$\epsup = 0.017$.

@@

We now introduce the theoretical framework that will be used to understand and quantify the experimental observations. This framework is based on a Galerkin truncation of the variable-depth Korteweg–de Vries (KdV) equation. The KdV equation is a commonly excepted model for describing the propagation of unidirectional, shallow-water waves, accounting for weak nonlinearity and weak dispersion over long timescales and large spatial scales. We will perform Galerkin truncation of this system at wavenumber $\Lambda = $8--32 to obtain a finite-dimensional dynamical system that exhibits weak turbulence. We outline the Hamiltonian structure of both the continuous KdV and the truncated system. This structure is exploited to obtain invariant measures of the underlying dynamics and to ultimately understand the experimental findings on anomalous wave statistics triggered by an ADC.

@@
